Arrows for Parallel Computations

### 2017-08-20 paper/content/benchmarks.tex
renamed figs (sudoku and torus)
filling in blanks in the benchmark text. now only the description for the benchmarks is needed for this chapter

\newcommand{\performanceSkelRMSM}[2]{
\performanceplot{Parallel run time of \rmtest ``#2''}{Eden CP, Multicore, Par Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-eden-cp-#1-#2.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-mult-#1-#2.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-par-#1-#2.csv};
}{17}{\plotwidthSMP}
}

\newcommand{\speedupSkelRMSM}[2]{
\speedupplot{Speedup of \rmtest ``#2''}{Eden CP, Multicore, Par Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-eden-cp-#1-#2.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-mult-#1-#2.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-par-#1-#2.csv};
}{17}{\plotwidthSMP}
}

@@

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupTorusDist{1024}{32}{260}\label{123123123}}%
%\hfill%
{\performanceTorusDistDiff{1024}{32}{0.5}{260}\label{12312312312}}
\caption{Parallel performance of \torustest on the Glasgow grid
  consisting of 256 cores. Input was a matrix size of $1024$. The top plot shows absolute speedup in a distributed memory setting. The
  closer to linear speedup the better. Time
  (and hence speedup) measurements for PArrows with Eden backend and
  Eden almost coincide. Hence, bottom plot shows
absolute time differences for this benchmark. The
lower the value, the better for PArrows\olcomment{CHECKME}.}
\label{123123123122}
\end{figure}

\newcommand{\performanceSudokuSM}[1]{
\performanceplot{Parallel run time of \sudokutest ``#1''}{Eden CP, Multicore, Par Monad, Par Monad (vanilla)}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-eden-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-mult-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-par-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parmonad-sudoku-sudoku17.#1.txt.csv};
}{17}{\plotwidthSMP}
}

\newcommand{\speedupSudokuSM}[1]{
\speedupplot{Speedup of \sudokutest ``#1''}{Eden CP, Multicore, Par Monad, Par Monad (vanilla)}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-eden-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-mult-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-par-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parmonad-sudoku-sudoku17.#1.txt.csv};
}{17}{\plotwidthSMP}
}

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupSudokuSM{1000}\label{1231231232}}%
%\hfill%
{\performanceSudokuSM{1000}\label{1231231123122}}
\caption{Relative speedup of \rmtest on a multicore machine. We used the same PArrows-based implementation with
  different backends on the same hardware and the |parMap| version from the |Par| Monad examples \citet{par-monad}\mbcomment{\\originally from http://community.haskell.org/~simonmar/par-tutorial.pdf}. Measurements were performed on a single node of the Glasgow
  grid; it has 16 real cores. Input was a file of $1000$ Sudokus. The
  closer to linear speedup the better.}
\label{1231231233122}
\end{figure}

@@

In the distributed memory setting the same effect ensues. We obtain
plummeting speedup of 124.31 at 512 hyperthreaded cores, whereas it was
213.172 for 256 real cores. Apparently, hyperthreading in the Glasgow
grid fails to execute two parallel Haskell processes with full-fledged
parallelism. For this reason, we did not regard hyperthreaded cores in
our speedup plots in Figs.~\ref{fig:bench-rm-sm}--XXXXXXXX.

@@

The difference between, say, PArrows with |Par| Monad backend and a
genuine |Par|
Monad benchmark is very small. To give an example, it is XXXXXXXX for
XXXXXXXXX and XXXXXXXXXX in the shared memory setting. It is almost invisible in speedup and
(non shown) run time plots. We thus show only the results for the
PArrows-enabled versions.

To showcase that the induced overhead of
PArrows is small, we plot execution time differences between measurements for
PArrows and the corresponding backend in a separate plot
(Figs.~\ref{fig:bench-rm-dist}, XXXXXXX). The differences range in
about 0.5~seconds for the execution time of 46~seconds on 256 cores
for distributed \rmtest with PArrows and Eden. The plots show absolute
time differences that are not relativated \wrt the total execution time.

>>>>

\newcommand{\performanceSkelRMSM}[2]{
\performanceplot{Parallel run time of \rmtest ``#2''}{Eden CP, Multicore, |Par| Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-eden-cp-#1-#2.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-mult-#1-#2.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-par-#1-#2.csv};
}{17}{\plotwidthSMP}
}

\newcommand{\speedupSkelRMSM}[2]{
\speedupplot{Speedup of \rmtest ``#2''}{Eden CP, Multicore, |Par| Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-eden-cp-#1-#2.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-mult-#1-#2.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-par-#1-#2.csv};
}{17}{\plotwidthSMP}
}

@@

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupTorusDist{1024}{32}{260}\label{subfig:speedupTorusDist}}%
%\hfill%
{\performanceTorusDistDiff{1024}{32}{0.5}{260}\label{subfig:performancetorusDistDiff}}
\caption{Parallel performance of \torustest on the Glasgow grid
  consisting of 256 cores. Input was a matrix size of $1024$. The top plot shows absolute speedup in a distributed memory setting. The
  closer to linear speedup the better. Time
  (and hence speedup) measurements for PArrows with Eden backend and
  Eden almost coincide. Hence, bottom plot shows
absolute time differences for this benchmark. The
lower the value, the better for PArrows\olcomment{CHECKME}.}
\label{fig:torusBenchmark}
\end{figure}

\newcommand{\performanceSudokuSM}[1]{
\performanceplot{Parallel run time of \sudokutest ``#1''}{Eden CP, Multicore, |Par| Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-eden-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-mult-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-par-sudoku17.#1.txt.csv};
%\addplot+ [very thick] table [scatter, x="nCores", y="time", col sep=comma, mark=none,
%smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parmonad-sudoku-sudoku17.#1.txt.csv};
}{17}{\plotwidthSMP}
}

\newcommand{\speedupSudokuSM}[1]{
\speedupplot{Speedup of \sudokutest ``#1''}{Eden CP, Multicore, |Par| Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-eden-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-mult-sudoku17.#1.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-par-sudoku17.#1.txt.csv};
%\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
%smooth]{benchmarks/sudoku-sm/bench-sudoku-sm.bench.parmonad-sudoku-sudoku17.#1.txt.csv};
}{17}{\plotwidthSMP}
}

\begin{figure}
\centering
%\performanceSkelRMDist{44497}{256}{32,64,128,256,512}{544}
%
{\speedupSudokuSM{1000}\label{subfig:speedupSudokuSM}}%
%\hfill%
{\performanceSudokuSM{1000}\label{subfig:performanceSudokuSM}}
\caption{Relative speedup of \rmtest on a multicore machine. We used the same PArrows-based implementation with
  different backends on the same hardware and the |parMap| version from the |Par| Monad examples. Measurements were performed on a single node of the Glasgow
  grid; it has 16 real cores. Input was a file of $1000$ Sudokus. The
  closer to linear speedup the better.}
\label{fig:sudokuSMBenchmark}
\end{figure}

@@

In the distributed memory setting the same effect ensues. We obtain
plummeting speedup of 124.31 at 512 hyperthreaded cores, whereas it was
213.172 for 256 real cores. Apparently, hyperthreading in the Glasgow
grid fails to execute two parallel Haskell processes with full-fledged
parallelism. For this reason, we did not regard hyperthreaded cores in
our speedup plots in Figs.~\ref{fig:bench-rm-sm}-\ref{fig:sudokuSMBenchmark}.

@@

The difference between, say, PArrows with |Par| Monad backend and a
genuine |Par|
Monad benchmark is very small. To give an example, it is ~0.4s in favour of PArrows for 16 cores (10.8s vs. 11.2s) and ~-0.8s in favour of the |Par| monad for 8 cores (16.1s vs. 16.9s) for
the Sudoku benchmark in the shared memory setting. It is almost invisible in speedup and
(non shown) run time plots. We thus show only the results for the
PArrows-enabled versions.

To showcase that the induced overhead of
PArrows is small, we plot execution time differences between measurements for
PArrows and the corresponding backend in a separate plot
(Figs.~\ref{fig:bench-rm-dist}, \ref{fig:torusBenchmark}). The differences range in
about 0.5~seconds for the execution time of 46~seconds on 256 cores
for distributed \rmtest with PArrows and Eden. The plots show absolute
time differences that are not relativated \wrt the total execution time.
